---
title: "Function Reference"
subtitle: "Complete guide to rePredRet functions"
---

## Overview

rePredRet provides functions for three main use cases:

1. **Query pre-built predictions** - Access predictions from the published model repository
2. **Build custom models** - Use your own RT data to build models
3. **Explore the database** - Search compounds and systems

---

## Querying Pre-built Predictions

These functions access the published rePredRet-data repository.

### `rePredRet_download()`

Downloads the latest predictions and models from GitHub.

```{r}
#| eval: false
# Download to default cache location
rePredRet_download()

# Force re-download
rePredRet_download(force_download = TRUE)

# Download to specific location
rePredRet_download(cache_dir = "my_data")
```

### `rePredRet_systems()`

Lists all available chromatographic systems.
```{r}
#| eval: false
systems <- rePredRet_systems()
head(systems)
#   id           name method.type n_compounds
# 1 0001      FEM_short          RP          85
# 2 0002       FEM_long          RP         112
```

### `rePredRet_models()`

Lists all available models with quality metrics.

```{r}
#| eval: false
models <- rePredRet_models()
head(models)
#   from_id to_id n_compounds median_error median_ci_width
# 1    0001  0002          42        0.15            0.28
```

### `rePredRet_predict()`

Gets predictions for a specific compound.

```{r}
#| eval: false
# Search by InChI
preds <- rePredRet_predict("InChI=1S/C8H10N4O2/c1-10-4-9-6-5(10)7(13)12(3)8(14)11(6)2/h4H,1-3H3")

# Search by name (if indexed)
preds <- rePredRet_predict("caffeine")
```

### `plot_model()`

Plots a pre-built model showing the calibration curve and confidence bands.

```{r}
#| eval: false
# Static ggplot
plot_model("0001", "0002")

# Interactive plotly
plot_model("0001", "0002", interactive = TRUE)

# Without confidence bands
plot_model("0001", "0002", show_ci = FALSE)

# Without calibration points
plot_model("0001", "0002", show_calibration = FALSE)
```

---

## Building Custom Models

Use your own RT data to build models and generate predictions.

### `build_user_models()`
Builds calibration models from RepoRT systems to your system.

```{r}
#| eval: false
models <- build_user_models(
  inchis = my_inchis,           # Character vector of InChI strings
  rts = my_rts,                 # Numeric vector of retention times
  system_type = "RP",           # "RP" or "HILIC"
  min_compounds = 10,           # Minimum compounds for model building
  method = "fast_ci",           # "fast_ci" (fast) or "bootstrap" (accurate)
  verbose = TRUE
)
```

**Returns:** `UserRTModels` S4 object with:

- `models`: Named list of model objects
- `diagnostics`: Per-model quality metrics
- `input_data`: Your original input
- `metadata`: Run information

**Methods:**

```{r}
#| eval: false
show(models)                    # Print summary
summary(models)                 # Detailed statistics
plot(models)                    # Diagnostic plots
plot(models, type = "calibration")  # Calibration curves
names(models)                   # List system IDs
models[["0001"]]               # Extract specific model
saveRDS(models, "models.rds")  # Save for later
```

### `predict_rt_user()`

Generates predictions using built models.

```{r}
#| eval: false
predictions <- predict_rt_user(models)

# Predict specific compounds only
predictions <- predict_rt_user(
  models,
  compounds = c("InChI=1S/C8H10N4O2/...")
)
```

**Returns:** `UserRTPredictions` S4 object with:

- `predictions`: Data frame of predictions (one per compound, best CI selected)
- `source_models`: System IDs used
- `metadata`: Run information

**Methods:**

```{r}
#| eval: false
show(predictions)               # Print summary
preds_df <- as.data.frame(predictions)  # Extract data frame
plot(predictions)               # CI width distribution
plot(predictions, type = "by_system")   # Predictions per system
```

### `compare_systems()`

Ranks repository systems by similarity to your system.

```{r}
#| eval: false
# Rank by R-squared (default)
compare_systems(models)

# Rank by RMSE
compare_systems(models, metric = "rmse")

# Top 5 most similar systems
compare_systems(models, metric = "r_squared", top_n = 5)
```

---

## Exploring the Database

### `search_compound()`

Searches for a compound across all systems.

```{r}
#| eval: false
# Search by name (partial match)
search_compound("caffeine")
search_compound("quercetin")

# Search by InChI
search_compound("InChI=1S/C8H10N4O2")

# Filter to RP systems only
search_compound("epicatechin", system_type = "RP")
```

**Returns:** Data frame with columns:

- `system_id`, `system_name`, `system_type`
- `compound_name`, `rt`, `inchi`

### `get_system_info()`

Gets detailed information about a chromatographic system.

```{r}
#| eval: false
info <- get_system_info("0001")
info$name           # System name
info$type           # RP or HILIC
info$n_compounds    # Number of compounds
info$metadata       # Column, gradient, etc.
info$gradient       # Gradient profile
```

---

## Data Loading (Advanced)

### `download_report()`

Downloads the RepoRT repository from GitHub.

```{r}
#| eval: false
# Download latest
path <- download_report()

# Download specific release
path <- download_report(release = "v1.0.0")
```

### `load_report_data()`

Loads datasets from a local RepoRT directory.

```{r}
#| eval: false
data <- load_report_data(path, method_types = c("RP", "HILIC"))
data$studies    # Study metadata
data$datasets   # List of datasets with RT data
```

### `get_common_compounds()`

Finds compounds present in two datasets.

```{r}
#| eval: false
common <- get_common_compounds(dataset1, dataset2)
# Returns matrix: [rt_source, rt_target]
```

---

## Summary Table

| Function | Purpose |
|----------|---------|
| `rePredRet_download()` | Download pre-built predictions |
| `rePredRet_systems()` | List available systems |
| `rePredRet_models()` | List available models |
| `rePredRet_predict()` | Query predictions for a compound |
| `plot_model()` | Plot a pre-built model |
| `build_user_models()` | Build models from your data |
| `predict_rt_user()` | Generate predictions |
| `compare_systems()` | Find similar systems |
| `search_compound()` | Search for compounds |
| `get_system_info()` | Get system details |
